---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(dplyr)
library(ggplot2)
library(scales)
library(ggstream)
options(warn = -1)
```

# ggstream

<!-- badges: start -->
[![R-CMD-check](https://github.com/davidsjoberg/ggstream/workflows/R-CMD-check/badge.svg)](https://github.com/davidsjoberg/ggstream/actions)
<!-- badges: end -->

The goal of `ggstream` is to create a simple but powerful implementation of streamplot/streamgraph in `ggplot2`. A streamplot is a stacked area plot mostly used for time series.

This is a development, expect breaking changes before submission to CRAN.

## Installation

You can install the development version of ggstream from github with:

``` r
remotes::install_github("davidsjoberg/ggstream")
```

## Examples

#### Type: Mirror

The characteristic streamplot which creates a symmetrical area chart around the x axis.

```{r pressure, echo = FALSE, fig.height=5, fig.width=9, dpi = 400, message=FALSE}
blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
    geom_stream(extra_span = 0.013, type = "mirror", n_grid = 3000, bw = .78) +
    geom_stream_label(size = 4, type = "mirror", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 18) +
  theme(legend.position = "none") +
  # scale_y_continuous(limits = c(-400, 400)) +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, billions")
```

#### Type: Ridge

Which is equivalent to a stacked area chart.

```{r pressure2, echo = FALSE, fig.height=5, fig.width=9, dpi = 400, message=FALSE}
blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
    geom_stream(extra_span = 0.013, type = "ridge", n_grid = 3000, bw = .78) +
    geom_stream_label(size = 4, type = "ridge", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 18) +
  theme(legend.position = "none") +
  # scale_y_continuous(limits = c(-400, 400)) +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, billions")
```


#### Type: Proportional

The type `proportional` shows the share of each group in percent. Stacked like the `ridge` type.

```{r pressure3, echo = FALSE, fig.height=5, fig.width=9, dpi = 400}
blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
  geom_stream(extra_span = 0.013, type = "proportional", n_grid = 3000, bw = .78) +
  geom_stream_label(size = 4, type = "proportional", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 18) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(1)) +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, percent")
```


## Basic usage

This is a basic example:

```{r example, fig.height=5, fig.width=9, dpi = 400}
library(ggstream)

ggplot(blockbusters, aes(year, box_office, fill = genre)) +
  geom_stream()
```


`ggstream` also features a custom labeling geom that places decent default labels.

```{r cars, fig.height=5, fig.width=9, dpi = 400}
ggplot(blockbusters, aes(year, box_office, fill = genre)) +
  geom_stream() +
  geom_stream_label(aes(label = genre))
```

#### Add some pimp

```{r cars8989, fig.height=5, fig.width=9, dpi = 400, message=FALSE}
library(cowplot)
library(paletteer)
library(dplyr)
library(colorspace)

blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
    geom_stream(extra_span = 0.013, type = "mirror", n_grid = 3000, bw = .78) +
    geom_stream_label(size = 4, type = "mirror", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 18) +
  theme(legend.position = "none") +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, billions")
```


## Advanced parameters

### bw

The main parameter to adjust in `geom_stream` is probably the bandwidth, or `bw`. A lower bandwidth creates a more bumpy plot and a higher bandwidth smoothens out some variation. Below is an illustration of how different bandwidths affect the stream plot.


```{r cars33, fig.height=7, fig.width=9, dpi = 400, message=FALSE}
library(patchwork)

base <- ggplot(blockbusters, aes(year, box_office, fill = genre)) + 
  theme(legend.position = "none")

(base +  geom_stream(bw = 0.5) + ggtitle("bw = 0.5")) /
(base +  geom_stream() + ggtitle("Default (bw = 0.75)")) /
(base +  geom_stream(bw = 1) + ggtitle("bw = 1"))
```


### Extra Span

Another important parameter to adjust is `extra_span`. This parameter adjust if a larger range than the range of the data which can help if the edges of the stream plot grows too large due in the estimation function. The additional range is set to y = 0 which forces the area towards zero. The cut-off can include the extra range or fit the data. Too illustrate this rather unintuitive parameter some variations are shown below. The transparent areas show the full estimation, and the solid area is the final plot.


```{r cars3344, fig.height=12, fig.width=9, dpi = 400, message=FALSE}
base <- ggplot(blockbusters, aes(year, box_office, fill = genre)) + 
  theme(legend.position = "none") +
  xlim(1970, 2028)

(base +  geom_stream() + ggtitle("Default")) /
(base +  geom_stream(extra_span = 0.001) + geom_stream(extra_span = 0.001, true_range = "none", alpha = .3) + ggtitle("extra_span = 0.001")) /
(base +  geom_stream(extra_span = .1) + geom_stream(extra_span = .1, true_range = "none", alpha = .3) + ggtitle("extra_span = .1")) /
(base +  geom_stream(extra_span = .2) + geom_stream(extra_span = .2, true_range = "none", alpha = .3) + ggtitle("extra_span = .2")) /
(base +  geom_stream(extra_span = .2, true_range = "none") + ggtitle("extra_span = .2 and true_range = \"none\""))
```

## Final remarks

The `ggstream` package provides some flexible ways to make stream plots but with decent defaults. However, due to the complexity of the underlying smoothing/estimation it should be used carefully and mostly for fun too illustrate major trends.

If you find a bug or have ideas for additional feature you are more than welcome to open an issue.
