---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
library(ggstream)
```

# ggstream

<!-- badges: start -->
<!-- badges: end -->

The goal of `ggstream` is to create a simple but powerful implementation of streamplot/streamgraph in `ggplot2`. A streamplot is a stacked area plot mostly used for time series.

This is a development, expect breaking changes before submission to CRAN.

## Installation

You can install the development version of ggstream from github with:

``` r
remotes::install_github("davidsjoberg/ggstream")
```

## Examples

The characteristic streamplot which creates a symmetrical area chart around the x axis.

```{r pressure, echo = FALSE, fig.height=5, fig.width=9, dpi = 400}
blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
    geom_stream(extra_span = 0.013, type = "mirror", n_grid = 3000, bw = .78) +
    geom_stream_label(size = 4, type = "mirror", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 24) +
  theme(legend.position = "none") +
  # scale_y_continuous(limits = c(-400, 400)) +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, billions")
```

Which is equivalent to a stacked area chart.

```{r pressure2, echo = FALSE, fig.height=5, fig.width=9, dpi = 400}
blockbusters %>%
  ggplot(aes(year, box_office, fill = genre, label = genre, color = genre)) +
    geom_stream(extra_span = 0.013, type = "ridge", n_grid = 3000, bw = .78) +
    geom_stream_label(size = 4, type = "ridge", n_grid = 1000) +
  cowplot::theme_minimal_vgrid(font_size = 24) +
  theme(legend.position = "none") +
  # scale_y_continuous(limits = c(-400, 400)) +
  scale_colour_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::darken(.8)) +
  scale_fill_manual(values = paletteer::paletteer_d("dutchmasters::milkmaid") %>% colorspace::lighten(.2)) +
  labs(title = "Box office per genre 1977-2019",
       x = NULL,
       y = "Current dollars, billions")
```

## Basic usage

This is a basic example:

```{r example, fig.height=5, fig.width=9, dpi = 400}
library(ggstream)

ggplot(blockbusters, aes(year, box_office, fill = genre)) +
  geom_stream()
```


`ggstream` also features a custom labeling geom that places decent default labels.

```{r cars, fig.height=5, fig.width=9, dpi = 400}
ggplot(blockbusters, aes(year, box_office, fill = genre)) +
  geom_stream() +
  geom_stream_label(aes(label = genre))
```


